---
title: "Joystick preprocessing"
---

Open joystick data .csv files. Note these files have no column names.
A trial is a row but different blocks are in different columns. The first column indicates the id number.

```{r}
library(tidyverse)
```

# Load trial information datafiles (acquired after runnning trialcode_addition.m on saloglo_input.mat)

```{r}

# TRIAL INFO

trl_info_path <- 'D:\\Dropbox\\MyScience\\MyPostdoc\\exp\\global_local\\2-data\\trial_info\\2-csv'

# select specific files inside folder
files = dir(path=trl_info_path, patter='trial_info')

trial_info <- files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(trl_info_path, .),  # the function read_csv() from the readr package
                 col_types = cols("i", "i", "i", "f", "f", "f", "f", "f", "c"), 
                 na = c("NaN"))) %>%
  reduce(rbind)  %>%      # reduce with rbind into one dataframe
  separate(trialcode, c("blocktype2", "salience", "lateness"), sep = c(1,2)) %>% #separate trialcode into three different columns
  arrange(id, block, trial) %>%
  rename(trial_in_block=trial) %>%
  select(-blocktype2) %>% #redundant column
  mutate(id = factor(id)) %>% 
  mutate(block = factor(case_when((block == 7  | block == 8)  ~ block-1,
                                  (block == 10 | block == 11) ~ block-2)))

   #  Trialcode
   #  % first digit: block type;
   #  % second digit: salience:
   #    % global standard is always 0;
   #    % in block type 1 (glostd=locdev):
   #    % X1=1 - change in ear;
   #    % Y2=2 - change in pitch;
   #  % third digit: count of lateness:
   #    % test block (block 1) and block with no deviance (block 12) are
   #    % excluded;
   #    % first 15 trials of each block is given 0;
   #    % each global deviant has value 0.

# Trial info is complete for every trial (in long format).
```

# Load joystick datafiles (after having been converted from .mat to .csv using mat_to_csv.m)

```{r}
# JOYSTICK DATA

joy_path <- 'D:\\Dropbox\\MyScience\\MyPostdoc\\exp\\global_local\\2-data\\joystick\\2-csv'

# select specific files inside folder
files = dir(path=joy_path, patter='joypos')

# read each selected file and combine into a unique dataframe
joy_data <- files %>%
  # read in all the files, appending the path before the filename
  map(~ read_csv(file.path(joy_path, .),  # the function read_csv() from the readr package
                 col_names = c("id", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15"), 
                 col_types = cols("i", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d", "d"), 
                 na = c("NaN", 0))) %>%
  reduce(rbind) %>%  # reduce with rbind into one dataframe

  # Note: ID 01 only has 15 columns, others have 16.
  # Joystick data is not yet in long format. Further pre-processing is required.
  transmute(id = id,
            `6` = `7`,
            `7` = `8`,
            `8` = `10`,
            `9` = `11`) %>%
  group_by(id) %>%
  mutate(trial_in_block = row_number()) %>%
  pivot_longer(cols = c(`6`,`7`,`8`,`9`), names_to = 'block', values_to = 'value') %>%
  drop_na() %>% # remove trials with value NA
  arrange(id, block) %>%
  mutate(trial = row_number()) %>%
  mutate(id = factor(id)) %>% 
  mutate(block = factor(block)) %>%
  mutate(block_type = case_when( (block == 6 | block == 7) ~ 'Default',
                                (block == 8 | block == 9) ~ 'Reversed')) %>%
  select(id, block, block_type, trial, trial_in_block, value) 

# Joystick data is now in long format. It can be merged with the trial info table.

df <- left_join(trial_info, joy_data, by = c('id', 'block', 'trial_in_block'))

# Further pre-processing is needed to normalize the joystick values.

```

## Joystick position values normalisation

Possible joystick values are between 0 and √2.
In blocks 6 and 7, participants were instructed that joystick(0)=prob(0) and 
joystick(√2)=prob(100).
In blocks 8 and 9, the values were reversed so joystick(√2)=prob(0) and joystick(0)=prob(100).

I will create a new variable `prob` that corrects for reversal, and another au
that normalizes for each block and id. In Yijie's report the denominator is said to be the max,
but in reality it is max-min.

```{r}
df <- df %>%
  mutate(prob = case_when((block_type == 'Default') ~ round((value/sqrt(2))*100, 2),
                          (block_type == 'Reversed') ~ round(100-((value/sqrt(2))*100),2))) %>%
  group_by(id, block) %>%
  mutate(min_prob = min(prob), max_prob = max(prob)) %>%
  mutate(au = round((prob-min_prob)/(max_prob-min_prob), 2)) %>%
  mutate(value = round(value, 2)) %>%
  ungroup() %>%
  select(-min_prob, -max_prob)
```

```{r}
out_fullfilename <- 'D:\\Dropbox\\MyScience\\MyPostdoc\\exp\\global_local\\2-data\\joystick\\3-preprocessed\\joystick.csv'
write_csv(df, path= out_fullfilename)
```


