---
title: "Joystick analysis"
---

# Load preprocessed dataset
```{r}
library(tidyverse)

#read from online file
df <- read_csv('https://www.dropbox.com/s/klfyv2hk78ko8a4/joystick.csv?dl=1',
         col_types = cols("f", "f", "i", "f", "f", "f", "f", "f", "i", "i", "f", "i", "d", "d", "d"))

#read from local drive
#df <- read_csv('D:\\Dropbox\\MyScience\\MyPostdoc\\exp\\global_local\\2-data\\joystick\\3-preprocessed\\joystick.csv',
#         col_types = cols("f", "f", "i", "f", "f", "f", "f", "f", "i", "i", "f", "i", "d", "d", "d"))
```

## Raw data visualisation

```{r}
df %>% 
  ggplot(aes(x=block, y=value)) +
    geom_jitter(size=.05, height = .0, width=.4) +
    facet_wrap(id~.)

df %>% 
  ggplot(aes(x=block, y=prob)) +
    geom_jitter(size=.05, height = .0, width=.4) +
    facet_wrap(id~.)

df %>% 
  ggplot(aes(x=block, y=au)) +
    geom_jitter(size=.05, height = .0, width=.4) +
    facet_wrap(id~.)
```

```{r eval=F}
df %>% 
  filter(lateness>0) %>%
  ggplot(aes(x=lateness, y=au)) +
    geom_jitter(size=.1, width=.4, height=.0) +
    geom_smooth(aes(group=id)) +
    facet_wrap(id~.)

df %>% 
  filter(lateness>0) %>%
  ggplot(aes(x=lateness, y=au)) +
    geom_jitter(size=.1, width=.4, height=.0) +
    geom_smooth() 
```


More preprocessing:
- Issue with id=12?
- add REAL likelihood of deviant to table?

###

- Remove unwanted trials (lateness = 0 : first 15 and globdev)
- Equalize number of trials?

## Analysis

How many points are not 0 or 1?

```{r}
df %>%
  group_by(id) %>%
  summarise(
    count_0 = sum(au<0.2),
    count_1 = sum(au>0.8),
    prop = round((count_0+count_1)/n(), 2)
  ) %>%
  arrange(prop)

```

# Linear mixed-model

```{r eval=F}

df_mdl %>%
  lmerTest::lmer(
    au ~ block + lateness + salience + (1|id), 
    data=.) %>%
  {. ->> mdl }

summary(mdl)

#coef(mdl)
#ranef(mdl)

mixed_model_df_fit <- df_mdl %>% 
  ungroup() %>%
  mutate(fit = predict(mdl))

mixed_model_df_fit %>%
  ggplot(aes(x=lateness, y=au)) + 
    geom_hline(yintercept=0.5, linetype="dashed") +
    #geom_violin(aes(groupsize = .2, width = .2, height = .0, alpha = .4) +
    geom_violin(aes(group=lateness, fill=as.factor(block)), alpha=.4) +
    #stat_summary(fun.y=mean, geom="line", size=.6) +
    geom_smooth(aes(y=fit, group=id, colour=factor(id)), method = 'lm', se=FALSE, size=.5, alpha = .5) + 
    #facet_wrap(id ~ .) +
    theme_bw() + 
    theme(aspect.ratio = 1, legend.position = 'right') +
    NULL


plot(residuals(mdl))
hist(residuals(mdl))
qqnorm(residuals(mdl))
```




# Bayesian model with brms


```{r}
library(brms)

# Step 2. Subset data
df_mdl <- df %>%
  #filter(!id==12) %>%
  filter(lateness > 0) 

# Step 3. Plotting the data
ggplot(data  = df_mdl,
       aes(x = lateness,
           y = au))+
  geom_point(size = 1.2,
             alpha = .8,
             position = "jitter")+ # to add some random noise for plotting purposes
  #geom_smooth(aes(y=fit, group=id, colour=factor(id)), method = 'lm', se=FALSE, size=.5, alpha = .5) +
  theme_minimal()+
  labs(title = "Rating by lateness")
```

Makes use of a Hamiltonian Monte Carlo sampler algorithm (MCMC) to approximate the posterior (distribution)

https://www.rensvandeschoot.com/tutorials/brms-started/

```{r}
brm_model <- brm(au ~ 1 + (1 | id), 
                              data   = df_mdl, 
                              warmup = 100, 
                              iter   = 200, 
                              chains = 2, 
                              inits  = "random",
                              cores  = 2)  #the cores function tells STAN to make use of 2 CPU cores simultaneously instead of just 1.
```

