---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
set.seed(123)
```

```{r}
# folder containing individual datasets 
path <- '/Users/santiago/Dropbox/MyScience/MyPostdoc/exp/global_local/3-analysis/eeg/4-R/DATA'

# select specific files inside folder
files = dir(path=path, patter='cnv_slopes_P')

# read each selected file and combine into a unique dataframe
data <- files %>%
  # using cols_only() to select only a subset of columns from the CSV file
  map(~ read_csv(file.path(path, .), 
                 col_names = c("id", 
                               "block", 
                               "trial_in_block", 
                               "blocktype", 
                               "lateness",
                               "slope"), 
                 col_types = cols("f", 
                                  "f", 
                                  "i", 
                                  "f", 
                                  "i",
                                  "d"), 
                 na = c("NaN"))) %>%
  reduce(rbind)

# Separate trialcode into various columns (as is already the case in joystick_preprocessing.Rmd)

#data <- data %>%
#  #separate(trialcode, c("blocktype2", "salience", "lateness"), sep = c(1,2)) %>%
#  mutate(lateness = as.numeric(substr(trialcode, nchar(trialcode), nchar(trialcode)))) # keep only last digit

```

# Load electrode labels and neighbours

`elecinfo.csv` is the file with electrode labels and neighbours.
```{r}
# Load
elecinfo <- read_csv('/Users/santiago/Dropbox/MyScience/MyPostdoc/exp/global_local/3-analysis/eeg/4-R/elecinfo.csv', col_types = cols('c', 'c') ) 

# Add to data
data_with_elecs <- data %>%
  group_by(id, block, trial_in_block) %>%
  mutate(elec = elecinfo$labels,
         neighbours = elecinfo$neighbours)
```


# Merge with behavioral data 
Note: in cnv slopes there are trials missing (trial nr start at >15).

```{r}
# Load behavioral metadata (only blocks 6-7-8-9)
#read from online file
metadata <- read_csv('https://www.dropbox.com/s/klfyv2hk78ko8a4/joystick.csv?dl=1',
                     col_types = cols("f", "i", "i", "f", "f", "f", "f", "f", "i", "i", "f", "i", "d", "d", "d"))

# Correct for block number (need the same in both)
#data %>% summarise(block = unique(block))

# revert block number change during preprocessing so it matches again with slope data

metadata <- metadata %>%
  mutate(block = factor(case_when((block == 6  | block == 7)  ~ block + 1,
                                  (block == 8  | block == 9)  ~ block + 2))) %>%
  select(-salience, -blocktype)

```

Join by `trial_in_block`

```{r}
#lateness <- metadata %>% select(trial_in_block, lateness)
#rm(data)  #for memory issue

joined <- data_with_elecs %>%
  #filter(block < 6) %>%
  left_join(., metadata, by = c('id', 'block', 'trial_in_block', 'lateness'))  #nest_join #inner_join

# More clean up
rm(data, data_with_elecs, elecinfo, metadata)

# Remove lateness = 0
df <- joined %>%
  filter(lateness>0)
```

# Reduction of electrodes

The following method is for data-driven eletrode selection. We skip it if we use the electrodes from the cluster-permutation.

## Sampling of trials

First, for each id, we will select a random subset of trials (__ %) per block to calculate the
3 neighbouring electrodes with the strongest modulation of CNV slope with lateness.


```{r, eval=FALSE}

#How many trials per level for each id and each block
df %>% group_by(id, block, lateness) %>% summarise(n=n()/91) -> summary
#min is 12!!! for lateness 4

#Select trials
n_sel <- 10  #half will be lateness 1 and half lateness 4

# Create vector of chosen trials
# Make it different for each block and for each participant
id_list <- list()
block_list <- list()
for (this_id in 1:21){
  for (this_block in c(2:5, 7:8, 10:11, 13:16)){  #actuql block nrs in data
    # find idx of lateness 1 and lateness 4 in each block to balance quantities
    smaller_df <- df %>% ungroup() %>% filter(id==this_id, block==this_block)
    indices_1 <- smaller_df %>% filter(lateness==1) %>% summarise(unique = unique(trial_in_block)) %>% pull()
    indices_4 <- smaller_df %>% filter(lateness==4) %>% summarise(unique = unique(trial_in_block)) %>% pull()
    #n <- length(smaller_df %>% summarise(unique = unique(trial_in_block)))
    sampled_1 <- sample(x = indices_1, size = n_sel/2, replace = FALSE)
    sampled_4 <- sample(x = indices_4, size = n_sel/2, replace = FALSE)
    block_list[this_block] <- list(c(sampled_1, sampled_4))
  }
  id_list[this_id] <- list(block_list)
  print(this_id)
}

# Sample same amount of lateness 1 and lateness 4

# find idx of lateness 1 and lateness 4 in each block
# combine them
# use that as list instead of `x = 1:max_trial_index`

# Select the sampled trials in a new dataframe
df <- df %>%
  rowwise() %>%
  mutate(sel = case_when((trial_in_block %in% id_list[[as.numeric(as.character(id))]][[as.numeric(as.character(block))]]) ~ 'training',
                         TRUE ~ 'validation'))  # NOT WORKING

training <- df %>%
  filter(sel=='training')

validation <- df %>%
  filter(sel == 'validation')
```

## Select best electrodes

```{r, eval=FALSE}
# Calculate for each id, all trials/blocks collapsed:
# Difference between lateness 1 and 4 at each electrode
# Rank electrodes by effect size
topelecs <- training %>%
  filter(lateness %in% c(1,4)) %>%
  group_by(id, elec) %>%
  summarise(mean_1 = mean(slope[lateness==1]),
            mean_4 = mean(slope[lateness==4]),
            mean_diff = abs(mean_4 - mean_1)) %>% #abs(mean_4 - mean_1)) %>%
  arrange(id, -mean_diff) %>%
  group_by(id) %>%
  slice_head(n=3, .preserve = TRUE)  %>% # preserve id grouping
  select(id, elec)

 # Now take the validation data and only keep the top electrodes from the training data
validation_simple <- validation %>%
  semi_join(., topelecs, by = c('id', 'elec')) %>%
  group_by(id, block, trial_in_block, lateness) %>% #to maintain lateness info
  summarise(val = mean(slope))   # Reduce to one value per trial (instead of 3)

```

```{r}
# List of electrodes from the cluster (from Matlab)
cluster_elecs <- c('E5', 'E6', 'E7', 'E11', 'E12', 'E13', 'E106', 'E111', 'E112', 'E117', 'E118')
  
# Select only those electrodes
topelecs <- df %>%
  filter(elec %in% cluster_elecs) %>%
  group_by(id, block, trial_in_block, lateness) %>%
  summarise(cluster_slope = mean(slope, na.rm = T)) # Average across those electrodes

```

Model

```{r}
library(lmerTest)

# HYPOTHESIS 1.
mdl1 <- lmer(cluster_slope ~ lateness + (1+lateness|id), data = topelecs)
summary(mdl1)
#ranef(mdl)

topelecs %>%
  ggplot(aes(x=lateness, y=cluster_slope)) + 
    geom_smooth(method='lm', alpha=.5, aes(group=id, color=id)) +
  #geom_jitter(alpha=.1) +
   stat_summary(fun.data = mean_cl_boot, geom = 'line') + 
  stat_summary(fun.data = mean_cl_boot, geom = 'pointrange') +  
  #facet_grid(~block) +
  NULL

```

```{r}
#HYPOTHESIS 5a
# Part A vs Part B: does the effect of lateness increase in part B?

topelecs <- topelecs %>%
  mutate(part = case_when((block == 2 | block == 3 | block == 4 | block == 5) ~ 'A',
                          (block == 7 | block == 8 | block == 10 |block == 11) ~ 'B',
                          (block == 13 | block == 14 |block == 15 |block == 16) ~ 'C'))
                          
topelecs %>%
  filter(part %in% c('A', 'B')) %>%
  lmer(cluster_slope ~ lateness + part + (1|id), data = .) -> mdl5a1

topelecs %>%
  filter(part %in% c('A', 'B')) %>%
  lmer(cluster_slope ~ lateness * part + (1|id), data = .) -> mdl5a2

anova(mdl5a1, mdl5a2)

# CCL: NO DIFFERENCE BETWEEN PART A AND PART B

# HYPOTHESIS 5A CONTINUED
# Part A vs Part C: does the effect of lateness also increase, depending on
# participant's report?

topelecs %>%
  filter(part %in% c('A', 'C')) %>%
  lmer(cluster_slope ~ lateness + part + (1|id), data = .) -> mdl5a3

topelecs %>%
  filter(part %in% c('A', 'C')) %>%
  lmer(cluster_slope ~ lateness * part + (1|id), data = .) -> mdl5a4

anova(mdl5a3, mdl5a4)

# Add report instead of Part C:
#case_when part C and reported yes: attended,
#part C and reported no: not_attended,

topelecs <- topelecs %>%
  mutate(atte = case_when((id == 1) ~ 'n', (id == 2) ~ 'n', (id == 3) ~ 'y',
                          (id == 4) ~ '?', (id == 5) ~ '?', (id == 6) ~ 'n',
                          (id == 7) ~ 'n', (id == 8) ~ 'n', (id == 9) ~ 'n',
                          (id == 10) ~ 'n', (id == 11) ~ 'n', (id == 12) ~ '?',
                          (id == 13) ~ '?', (id == 14) ~ 'y', (id == 15) ~ 'y',
                          (id == 16) ~ 'n', (id == 17) ~ 'n', (id == 18) ~ 'n',
                          (id == 19) ~ 'y', (id == 20) ~ 'n', (id == 21) ~ 'y'))

mdl5a5 <- topelecs %>%
  filter(part %in% c('A', 'C')) %>%
  filter(atte %in% c('y', 'n')) %>%
  lmer(cluster_slope ~ lateness * part + (1|id), data = .)

mdl5a6 <- topelecs %>%
  filter(part %in% c('A', 'C')) %>%
  filter(atte %in% c('y')) %>%
  lmer(cluster_slope ~ lateness * part + (1|id), data = .)

#anova(mdl5a5, mdl5a6)

#CCL: NO INT EFFECT FOR ALL GROUP, STILL NO INT EFFECT BY ONLY KEEPING THOSE WHO REPORTED PREDICTING THE SEQUENCES


# Another try
mdl5a7 <- topelecs %>%
  lmer(cluster_slope ~ lateness * part * atte + (1|id), data = .)

summary(mdl5a7)

#CCL: block effect for Y in last block? (not a lateness interaction)
topelecs %>%
  ggplot(aes(x=lateness, y=cluster_slope, group=id, color=id)) + 
  #geom_jitter(alpha=.1) +
  #geom_smooth(method='lm', se=F) +
  facet_grid(atte~part) +
  stat_summary(fun.data = mean_cl_boot, geom = 'line') + 
  #stat_summary(fun.data = mean_cl_boot, geom = 'pointrange') +  
  NULL


```



```{r}
# New

# HYPOTHESIS 1.
# Electrodes as random variable
#df %>%
#  filter(elec %in% cluster_elecs) %>%
#  lmer(slope ~ lateness + (1+lateness|id/elec), data = .) %>%
#  summary()
#ranef(mdl)

# HYP 5
# New

topelecs2 <- df %>%
  filter(elec %in% cluster_elecs)

topelecs2 <- topelecs2 %>%
  mutate(part = case_when((block == 2 | block == 3 | block == 4 | block == 5) ~ 'A',
                          (block == 7 | block == 8 | block == 10 |block == 11) ~ 'B',
                          (block == 13 | block == 14 |block == 15 |block == 16) ~ 'C'))

topelecs2 <- topelecs2 %>%
  mutate(atte = case_when((id == 1) ~ 'n', (id == 2) ~ 'n', (id == 3) ~ 'y',
                          (id == 4) ~ '?', (id == 5) ~ '?', (id == 6) ~ 'n',
                          (id == 7) ~ 'n', (id == 8) ~ 'n', (id == 9) ~ 'n',
                          (id == 10) ~ 'n', (id == 11) ~ 'n', (id == 12) ~ '?',
                          (id == 13) ~ '?', (id == 14) ~ 'y', (id == 15) ~ 'y',
                          (id == 16) ~ 'n', (id == 17) ~ 'n', (id == 18) ~ 'n',
                          (id == 19) ~ 'y', (id == 20) ~ 'n', (id == 21) ~ 'y'))


topelecs2 %>%
  lmer(slope ~ lateness * part * atte + (1+lateness|id/elec), data = .) %>%
  summary()
```

